---
title: "Health disparity framework: Figure 2 and 3"
author: "Katherine N Thompson"
date: "2026-01-07"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    number_sections: false
    highlight: monochrome
    theme: flatly
    code_folding: show
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      comment = NA,
                      prompt = FALSE,
                      cache = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      results = 'markup')
options(bitmapType = 'quartz')
```

```{r clear global environment, include=FALSE}
remove(list = ls())
```

```{r set seed, include=FALSE}
set.seed(123)
```

```{r Load packages, include=FALSE}
library(knitr)
library(tidyverse)
library(viridis)
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggpattern)
library(cowplot)
library(grid)
library(dplyr)
```

```{r directories - EDITED CHUNK, include=FALSE}
# Directories updated to point at the simulated bootstrap results
# produced by the reviewer copies of pgs.R, race.R, ses.R, and sex.R,
# and to a local plots folder (instead of Katie's private paths). - EDITED LINE
results_dir <- "/Users/qtnzknt/Library/CloudStorage/OneDrive-UniversityCollegeLondon/Documents/healthdis/NatHB/code_review/simulated_data_scripts/data/confound/all/"  # EDITED LINE
plots_dir   <- "/Users/qtnzknt/Library/CloudStorage/OneDrive-UniversityCollegeLondon/Documents/healthdis/NatHB/code_review/simulated_data_scripts/plots/"  # EDITED LINE
dir.create(plots_dir, recursive = TRUE, showWarnings = FALSE)  # EDITED LINE
```

# README

The overall aim of this study is to understand the extent to which an intervention on social connection in childhood can mitigate the association between vulnerable characteristics and depression in adulthood. Using the health disparity framework, we will test mediation models between vulnerable characteristics (genetic vulnerability, socioeconomic position, sex, and race), social connection, and depression symptoms.

Our research questions are as follows:

1. How much of the disparity in depression symptoms (Y) between polygenic risk (X=0,1,2,3,4) quintiles would remain, if we were to intervene on inclusion, community, social connection, and friendship (M) so that those with the highest genetic risk receive the same support and inclusion as those with the lowest genetic risk (M(X=0)).
2. How much of the disparity in depression symptoms (Y) between males (X=0) and females (X=1) would remain, if we were to intervene on inclusion, community, social connection, and friendship (M) so that women receive the same support and inclusion as men (M(X=0)).
3. How much of the disparity in depression symptoms (Y) would remain both between White (X=0) and Black (X=1) people and White (X=0) and Asian (X=1) people, if we were to intervene on inclusion, community, social connection, and friendship (M) so that Black or Asian people receive the same support and inclusion as White people (M(X=0)).
4. How much of the disparity in depression symptoms (Y) between people from low (X=0) and high (X=1) socioeconomic position families would remain, if we were to intervene on inclusion, community, social connection, and friendship (M) so that low socioeconomic position families receive the same support and inclusion as high socioeconomic position (M(X=0)).

We adapt code from Herle et al. (2023), which applies a health disparity framework to evaluate whether an intervention on physical activity could mitigate the association between genetic liability for obesity (measured using a BMI polygenic score) and childhood/adolescent BMI.

Data preparation is done within the script *health_disparity_dataprep.Rmd* and analyses are described in *health_disparity_social.Rmd*

Description of the effects included in the plots:
* DE / IDM-DE / ME estimates the disparity that remains after intervening on the mediator (social connection), i.e. the remaining association between sex and depression symptoms if everyone had their social connection level shifted to match that of males. <!-- EDITED LINE: CDM renamed to ME -->
* TCE / Adj-TA is the total association between sex and depression, after adjusting for covariates.
* DIF reflects the potential mitigation due to intervening on social support and inclusion. Larger differences between Adj-TA and IDM-DE suggest greater scope for intervention.
* dep_1 and dep_0 are the same across mediators because they come from the total effect model (Y ~ sex + H1GI1Y), which does not include the mediator.

**Here, we create Figure 2 and 3 for the paper**

***

# Prep

```{r mediators}
# list of mediators
mediators <- c(
  "family",
  "school",
  "friends",
  "neigh",
  "religion",
  "social"
)

# Ns
n_dat_pheno_W2 = "9,987"
n_dat_pheno_W4 = "11,139"
n_dat_pheno_pgs_W2 = "4,478"
n_dat_pheno_pgs_W4 = "6,242"

comparison_levels <- c("SES: Low", "SES: Mid", "Sex: Female", "PGS: High", "PGS: Mid", "Race: Asian", "Race: Black")
mediator_levels <- c("Social","Family","School","Friends","Neigh","Religion")

poster.pal <- c(
  "Sex: Female"       = "#716f35",
  "PGS: High"      = "#814256",
  "PGS: Mid"    = "#CC79A7",
  "SES: Low"      = "#eccc6e",
  "SES: Mid"   = "#D55E00",
  "Race: Asian"       = "#a66940",
  "Race: Black"       = "#cfb991"
)

PAIR_ORDER <- c(
  "Sex: Female | Short-term (one year)",
  "Sex: Female | Long-term (one decade)",
  "PGS: High | Short-term (one year)",
  "PGS: High | Long-term (one decade)",
  "PGS: Mid | Short-term (one year)",
  "PGS: Mid | Long-term (one decade)",
  "SES: Low | Short-term (one year)",
  "SES: Low | Long-term (one decade)",
  "SES: Mid | Short-term (one year)",
  "SES: Mid | Long-term (one decade)",
  "Race: Asian | Short-term (one year)",
  "Race: Asian | Long-term (one decade)",
  "Race: Black | Short-term (one year)",
  "Race: Black | Long-term (one decade)"
)

comp_labels <- c(
  "Female vs Male"       = "Female",
  "High PGS vs Low"      = "High PGS",
  "Middle PGS vs Low"    = "Middle PGS",
  "Low SES vs High"      = "Low SES",
  "Middle SES vs High"   = "Middle SES",
  "Asian vs White"       = "Asian",
  "Black vs White"       = "Black"
)

WAVE_ORDER <- c("Short-term (one year)", "Long-term (one decade)")

MED_LABELS <- c(
  "Social"   = "Overall \nsocial",
  "Family"   = "Family \nsupport",
  "School"   = "School \nbelonging",
  "Friends"  = "Friendship \nconnections",
  "Neigh"    = "Neighbourhood \ncohesion",
  "Religion" = "Religious \ninvolvement"
)
```

# Results

Sex

```{r read in sens W24 sex results files}
results_c_all_sex_W2 <- readRDS(paste0(results_dir, "sex/bootstrap_results_W2.rds"))  # EDITED LINE
results_c_all_sex_W4 <- readRDS(paste0(results_dir, "sex/bootstrap_results_W4.rds"))  # EDITED LINE

# combine W2 and W4
results_c_all_sex <- rbind(results_c_all_sex_W2, results_c_all_sex_W4) %>% mutate(`Risk factor` = "Sex: Female")
```

SES

```{r read in sens c ses results files}
results_c_all_SESlow_W2 <- readRDS(paste0(results_dir, "ses/bootstrap_results_W2_SESlow.rds"))  # EDITED LINE
results_c_all_SESlow_W4 <- readRDS(paste0(results_dir, "ses/bootstrap_results_W4_SESlow.rds"))  # EDITED LINE
results_c_all_SESmid_W2 <- readRDS(paste0(results_dir, "ses/bootstrap_results_W2_SESmid.rds"))  # EDITED LINE
results_c_all_SESmid_W4 <- readRDS(paste0(results_dir, "ses/bootstrap_results_W4_SESmid.rds"))  # EDITED LINE

# combine W2 and W4
results_c_all_ses <- rbind(results_c_all_SESlow_W2, results_c_all_SESlow_W4, results_c_all_SESmid_W2, results_c_all_SESmid_W4) %>% rename(`Risk factor` = ses)
```

Race

```{r read in sens c race results files}
results_c_all_asian_W2 <- readRDS(paste0(results_dir, "race/bootstrap_results_W2_asian.rds"))  # EDITED LINE
results_c_all_black_W2 <- readRDS(paste0(results_dir, "race/bootstrap_results_W2_black.rds"))  # EDITED LINE
results_c_all_natam_W2 <- readRDS(paste0(results_dir, "race/bootstrap_results_W2_natam.rds"))  # EDITED LINE
results_c_all_asian_W4 <- readRDS(paste0(results_dir, "race/bootstrap_results_W4_asian.rds"))  # EDITED LINE
results_c_all_black_W4 <- readRDS(paste0(results_dir, "race/bootstrap_results_W4_black.rds"))  # EDITED LINE
results_c_all_natam_W4 <- readRDS(paste0(results_dir, "race/bootstrap_results_W4_natam.rds"))  # EDITED LINE

# combine W2 and W4
results_c_all_race <- rbind(results_c_all_asian_W2, results_c_all_black_W2, results_c_all_natam_W2,
                           results_c_all_asian_W4, results_c_all_black_W4, results_c_all_natam_W4) %>% rename(`Risk factor` = Race)
```

PGS

```{r read in sens c pgs results files}
results_c_all_pgs2_W2 <- readRDS(paste0(results_dir, "pgs/bootstrap_results_ALL_W2_pgs_reg2.rds"))  # EDITED LINE
results_c_all_pgs2_W4 <- readRDS(paste0(results_dir, "pgs/bootstrap_results_ALL_W4_pgs_reg2.rds"))  # EDITED LINE
results_c_all_pgs3_W2 <- readRDS(paste0(results_dir, "pgs/bootstrap_results_ALL_W2_pgs_reg3.rds"))  # EDITED LINE
results_c_all_pgs3_W4 <- readRDS(paste0(results_dir, "pgs/bootstrap_results_ALL_W4_pgs_reg3.rds"))  # EDITED LINE

# combine W2 and W4
results_c_all_pgs <- rbind(results_c_all_pgs2_W2, results_c_all_pgs2_W4, results_c_all_pgs3_W2, results_c_all_pgs3_W4) %>% rename(`Risk factor` = pgs)
```

## Join all

```{r join item results together}
results_c_all_sens <- rbind(
  results_c_all_sex,
  results_c_all_ses,
  results_c_all_race,
  results_c_all_pgs
) %>%
  mutate(`Risk factor` = factor(
    case_when(
      `Risk factor` == "SESlow" ~ "SES: Low",
      `Risk factor` == "SESmid" ~ "SES: Mid",
      `Risk factor` == "Sex: Female" ~ "Sex: Female",
      `Risk factor` == "pgs_reg3" ~ "PGS: High",
      `Risk factor` == "pgs_reg2" ~ "PGS: Mid",
      `Risk factor` == "asian" ~ "Race: Asian",
      `Risk factor` == "black" ~ "Race: Black"
      ))
  )
```

# Data prep

```{r plot spacing}
K    <- length(comparison_levels)
band <- 0.90                    # total width of each mediator “bundle”
step <- band / K
offset_seq <- seq(-(K-1)/2, (K-1)/2) * step
offsets    <- setNames(rev(offset_seq), comparison_levels)
sep_x <- seq(1.5, length(mediator_levels) - 0.5, by = 1)
axis_min <- -20; axis_max <- 35
band_pair   <- 0.95
step_pair   <- band_pair / length(PAIR_ORDER)
pair_offset <- setNames(seq(-(length(PAIR_ORDER)-1)/2, (length(PAIR_ORDER)-1)/2) * step_pair, PAIR_ORDER)

## Comparison order index
comp_order <- tibble::tibble(
  `Risk factor`   = comparison_levels,
  order_index  = seq_along(comparison_levels)
)
```

```{r format results data}
plot_results_c_all_sens <- results_c_all_sens %>%
  group_by(Wave, `Risk factor`) %>%
  filter(Effect_Type %in% c("ME", "TCE")) %>%  # EDITED LINE (was: CDM)
  dplyr::select(`Risk factor`, Mediator, Wave, Effect_Type, Mean) %>%
  pivot_wider(names_from = Effect_Type, values_from = Mean) %>%
  left_join(
    results_c_all_sens %>%
      filter(Effect_Type == "DIF") %>%
      dplyr::select(`Risk factor`, Mediator, Wave, DIF = Mean, CI_Low, CI_High), # confidence intervals are for DIF
    by = c("Mediator", "Wave", "Risk factor")
  ) %>%
  mutate(
    Significant = if_else(`CI_Low` > 0 | `CI_High` < 0, "Yes", "No"),
    Wave        = trimws(Wave),
    Wave_label  = recode(Wave,
                         "W2" = "Short-term (one year)",
                         "W4" = "Long-term (one decade)"),
    Wave_label  = factor(Wave_label,
                         levels = c("Short-term (one year)",
                                    "Long-term (one decade)")),
    Mediator = str_to_sentence(Mediator),
    Mediator    = factor(Mediator, levels = mediator_levels),
    Sig         = Significant == "Yes",
    `Percent difference` = (TCE - ME) / TCE * 100,  # EDITED LINE (was: CDM)
    Sig_position = `Percent difference` + 5,
    med_idx = as.numeric(factor(Mediator, levels = mediator_levels)),
    x_pos   = med_idx + offsets[as.character(`Risk factor`)], # comparison-specific fixed offset inside each mediator bundle
    Pair        = factor(paste(`Risk factor`, Wave_label, sep = " | "),
                         levels = PAIR_ORDER),
    med_idx     = as.integer(Mediator),
    x_adj       = med_idx + as.numeric(pair_offset[as.character(Pair)]),
    Worsened = case_when(
    `Percent difference` < 0 & Sig == FALSE ~ FALSE,
    `Percent difference` < 0 & Sig == TRUE ~ TRUE,
    `Percent difference` > 0 & Sig == FALSE ~ FALSE,
    `Percent difference` > 0 & Sig == TRUE ~ FALSE,
    TRUE ~ FALSE)
  ) %>%
  ungroup()
```

```{r arrow specific data prep}
## Use unique Mediator x Comparison combos to define x_base
order_df <- plot_results_c_all_sens %>%
  filter(`Risk factor` %in% comparison_levels) %>%
  distinct(Mediator, `Risk factor`) %>%
  left_join(comp_order, by = "Risk factor") %>%
  group_by(Mediator) %>%
  arrange(order_index, .by_group = TRUE) %>%
  mutate(
    rank_within = dplyr::row_number(),
    n_c_allomp      = dplyr::n(),
    med_index   = match(Mediator, mediator_levels),
    # narrower spread (0.12) keeps all points between separator lines
    x_base      = med_index + (rank_within - (n_c_allomp + 1)/2) * 0.12
  ) %>%
  ungroup()

## Join x_base, compute wave-specific x_plot
plot_results_c_all_sens <- plot_results_c_all_sens %>%
  filter(`Risk factor` %in% comparison_levels) %>%
  select(-dplyr::any_of(c("x_base", "x_plot"))) %>%
  left_join(order_df %>% select(Mediator, `Risk factor`, x_base),
            by = c("Mediator", "Risk factor")) %>%
  mutate(
    x_plot = dplyr::case_when(
      Wave_label == "Short-term (one year)"   ~ x_base - 0.025,
      Wave_label == "Long-term (one decade)"  ~ x_base + 0.025,
      TRUE                                    ~ x_base
    )
  )

tol <- 1e-6  # or 1e-4 if you want a looser threshold

arrow_df <- plot_results_c_all_sens %>%
  filter(
    !is.na(TCE),
    !is.na(ME),  # EDITED LINE (was: CDM)
    abs(TCE - ME) > tol   # only draw arrows when there is a visible change  # EDITED LINE (was: CDM)
  )

## Arrow style (single solid style)
dir_arrow <- grid::arrow(
  type   = "open",
  length = unit(5, "pt"),
  angle  = 25
)
```

# Figure 2

```{r plot 2 short with arrows}
# Dynamic y-axis limits (data-driven, not hardcoded) so no risk factor's
# TCE/ME estimate is silently dropped by an axis range tuned for different
# data. - EDITED LINE
y_range_fig2 <- range(c(plot_results_c_all_sens$TCE, plot_results_c_all_sens$ME), na.rm = TRUE)  # EDITED LINE
y_limits_fig2 <- c(min(0, y_range_fig2[1]), y_range_fig2[2] * 1.08)  # EDITED LINE (was: hardcoded c(0, 0.4))
y_breaks_fig2 <- scales::pretty_breaks(n = 5)(y_limits_fig2)  # EDITED LINE (was: hardcoded seq(0, 0.4, 0.1))

p_c_all_arrow <- ggplot() +
  geom_vline(xintercept = sep_x, colour = "grey39", linewidth = 0.2) +

  # --- Solid arrows for ALL effects
  geom_segment(
    data = arrow_df,
    aes(
      x     = x_plot,
      xend  = x_plot,
      y     = TCE,
      yend  = ME,  # EDITED LINE (was: CDM)
      color = `Risk factor`
    ),
    linewidth   = 0.6,
    lineend     = "round",
    arrow       = dir_arrow,
    show.legend = FALSE,
    na.rm       = TRUE
  ) +

  # --- TCE points
  geom_point(
    data = plot_results_c_all_sens,
    aes(
      x     = x_plot,
      y     = TCE,
      color = `Risk factor`,
      shape = Wave_label
    ),
    size   = 4,
    stroke = 0,
    na.rm  = TRUE
  ) +

  # --- Comparison colours (points only)
  scale_color_manual(
    values = poster.pal,
    limits = comparison_levels,
    labels = comp_labels[comparison_levels],
    drop   = FALSE,
    name   = NULL,
    guide  = guide_legend(
      order = 1,
      override.aes = list(
        shape    = 16,
        size     = 4,
        linetype = 0
      )
    )
  ) +

  # --- Short vs long term shapes
  scale_shape_manual(
    name   = "",
    values = c(
      "Short-term (one year)"   = 15,  # filled square
      "Long-term (one decade)"  = 16   # filled circle
    ),
    breaks = WAVE_ORDER,
    labels = WAVE_ORDER,
    guide  = guide_legend(
      order = 2,
      override.aes = list(
        color = "grey30",
        size  = 4
      )
    )
  ) +

  # --- Axes: all points stay between category separators
  scale_x_continuous(
    breaks = seq_along(mediator_levels),
    labels = unname(MED_LABELS[mediator_levels]),
    expand = expansion(add = 0.25)
  ) +
  scale_y_continuous(
    limits = y_limits_fig2,  # EDITED LINE (was: c(0, 0.4))
    breaks = y_breaks_fig2,  # EDITED LINE (was: seq(0, 0.4, 0.1))
    labels = scales::number_format(accuracy = 0.01),
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  labs(
    y = "Estimate",
    x = NULL
  ) +

  theme_classic(base_size = 18) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(color = "grey85", linewidth = 0.35),
    legend.position    = "bottom",
    legend.direction   = "horizontal",
    legend.box         = "horizontal",
    legend.box.just    = "left",
    legend.key.height  = unit(10, "pt"),
    legend.spacing.y   = unit(0.5, "pt"),
    legend.spacing.x   = unit(0.05, "pt"),
    axis.title.y       = element_text(size = 15, margin = margin(r = 14)),
    legend.text        = element_text(size = 12),
    plot.margin        = margin(6, 8, 6, 6)
  )

# EDITED LINE (output path)
ggsave(paste0(plots_dir, "Figure_2.png"),
        p_c_all_arrow, width = 12, height = 6, units = "in", dpi = 300)
```

# Figure 3

Option A: scaled % plot

```{r flip % sign}
# flip sign of % for scaled version of bar graph
plot_results_c_all_sens <- plot_results_c_all_sens %>%
  mutate(
    `Percent difference_flip` =
      `Percent difference` * -1
  )
```


```{r scaled % plot}
# positioning
ylims_pct <- c(-30, 70)
me_max <- max(abs(plot_results_c_all_sens$ME), na.rm = TRUE)  # EDITED LINE (was: cdm_max/CDM)
k <- 0.8 * max(abs(ylims_pct)) / me_max  # EDITED LINE (was: cdm_max)
pd <- position_dodge(width = step * 0.9)
ymax <- max(plot_results_c_all_sens$TCE, na.rm = TRUE)
ylims <- c(
  min(0, min(c(plot_results_c_all_sens$TCE, plot_results_c_all_sens$ME), na.rm = TRUE)),
  max(c(plot_results_c_all_sens$TCE, plot_results_c_all_sens$ME), na.rm = TRUE) * 1.15  # extra headroom for % labels above bars
)  # EDITED LINE (was: hardcoded c(0, 0.4))
bar_w <- step * 0.85

# plot
p_c_all_scaled <- ggplot(
  plot_results_c_all_sens,
  aes(
    x = x_pos,
    group = interaction(`Risk factor`, Wave_label)
    )
  ) +
  # 1) Total effect bar
  geom_col(
    aes(y = TCE,
        fill = `Risk factor`),
    width = bar_w,
    position = pd,
    color = "white",
    show.legend = c(fill = TRUE)
  ) +

  # 2) Mitigated effect bar
  geom_col_pattern(
    aes(
      y = ME,  # EDITED LINE (was: CDM)
      fill = `Risk factor`,
      fill = after_scale(scales::col_mix(fill, "white", 0.50)),
      pattern = if_else(Sig, "sig", "ns"),
      colour  = Worsened
    ),
    width = bar_w,
    position  = pd,
    pattern_colour = "white",
    pattern_fill   = NA,
    pattern_angle  = 45,
    pattern_density = 0.15,
    pattern_spacing = 0.01,
    pattern_size    = 0.3,
    show.legend = c(pattern = TRUE, fill = FALSE)
  ) +

  # 3) % reduction label
  geom_text(
    aes(
      y = TCE,
      label = if_else(
      Sig,
      paste0(round(`Percent difference_flip`, 0), "%"),
      NA_character_
    ),
      vjust = if_else(plot_results_c_all_sens$Worsened, -4.4, -0.7)
    ),
    position = pd,
    size = 3,
    show.legend = FALSE,
    fontface = "bold"
  ) +

  geom_vline(xintercept = sep_x, colour = "grey85", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.4, color = "grey45") +

  scale_x_continuous(
    breaks = seq_along(mediator_levels),
    labels = MED_LABELS
  ) +

  facet_grid(Wave_label ~ .) +

  scale_fill_manual(
    values = poster.pal,
    limits = comparison_levels,
    drop   = FALSE
  ) +

  scale_pattern_manual(
    name   = "",
    values = c(sig = "none", ns = "stripe"),
    labels = c(sig = "Significant", ns = "Not significant"),
    guide  = guide_legend(
      order = 2,
      override.aes = list(fill = "grey60", colour = "white")
    )
  ) +

  scale_y_continuous(
    limits = ylims,
    expand = expansion(mult = c(0.02, 0.10))  # extra top room for % labels
  ) +

  scale_colour_manual(
  values = c(
    `TRUE`  = "#2B1700",  # brown outline when worsened # A8422D
    `FALSE` = "white"     # white outline otherwise
  ),
  guide = "none"
  ) +

  labs(
    x = NULL,
    y = "Effect size",
    fill = NULL
  ) +

  theme_classic(base_size = 20) +
  theme(
    panel.grid.major.x = element_line(color = "grey85", size = 0.35),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.background   = element_blank(),
    strip.text         = element_text(size = 14, face = "bold"),
    axis.text          = element_text(color = "black"),
    axis.title.x       = element_text(size = 16, margin = margin(t = 6)),
    legend.position    = "bottom",
    legend.direction   = "horizontal",
    legend.key.height  = unit(10, "pt"),
    legend.key.width   = unit(20, "pt"),
    legend.text        = element_text(size = 16),
    plot.margin        = margin(6, 8, 6, 6)
  ) +
  guides(
    fill = guide_legend(
      order = 1,
      nrow = 1, byrow = TRUE,
      override.aes = list(
        pattern = "none",
        pattern_colour = NA,
        pattern_fill   = NA
      )
    )
  )

# EDITED LINE (output path)
ggsave(paste0(plots_dir, "Figure_3_PB_scaled.png"), p_c_all_scaled, width = 16, height = 8, units = "in", dpi = 300)
```

```{r bar chart of percentages}
# Dynamic y-axis limits for the percent-difference chart (data-driven, not
# hardcoded), so no risk factor is silently dropped. Percent difference can
# be numerically unstable when TCE is close to zero (division blow-up), so
# only finite values are used to set the range. - EDITED LINE
pct_finite   <- plot_results_c_all_sens$`Percent difference`[is.finite(plot_results_c_all_sens$`Percent difference`)]  # EDITED LINE
pct_range    <- range(pct_finite, na.rm = TRUE)  # EDITED LINE
pct_pad      <- diff(pct_range) * 0.15  # EDITED LINE
pct_limits   <- c(pct_range[1] - pct_pad, pct_range[2] + pct_pad)  # EDITED LINE (was: hardcoded c(-20, 50))
pct_breaks   <- scales::pretty_breaks(n = 6)(pct_limits)  # EDITED LINE (was: hardcoded seq(-20, 60, 10))

p_c_all <- ggplot(
  plot_results_c_all_sens,
  aes(
    x = x_pos,
    y = `Percent difference`,
    fill = `Risk factor`,
    group = `Risk factor`)
  ) +
  geom_col_pattern(
    aes(pattern = if_else(Sig, "sig", "ns")),
    width = step * 0.92,
    color = "white",
    pattern_colour = "white",
    pattern_fill   = NA,
    pattern_angle  = 45,
    pattern_density = 0.15,
    pattern_spacing = 0.01,
    pattern_size    = 0.3,
    show.legend = c(pattern = TRUE, fill = TRUE)
  ) +
  geom_text(
    aes(
      y = `Percent difference`,
      label = if_else(
      Sig,
      paste0(round(`Percent difference`, 0), "%"),
      NA_character_
    ),
      vjust = if_else(plot_results_c_all_sens$Worsened, 2, -0.7)
    ),
    size = 3,
    show.legend = FALSE,
    fontface = "bold"
  ) +
  geom_vline(xintercept = sep_x, colour = "grey85", size = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.4, color = "grey45") +
  scale_x_continuous(
    breaks = seq_along(mediator_levels),
    labels = MED_LABELS
  ) +
  # coord_flip(clip = "off") +
  # facet_wrap(~ Wave_label, nrow = 1) +
  facet_grid(Wave_label ~ .) +

  scale_fill_manual(
  values = poster.pal,
  limits = comparison_levels,   # keeps order
  drop   = FALSE) +
  scale_pattern_manual(
  name   = "",
  values = c(sig = "none", ns = "stripe"),
  labels = c(sig = "Significant", ns = "Not significant"),
  guide  = guide_legend(
    order = 2,
    override.aes = list(
      fill   = "grey60",  # neutral fill so stripes are clear
      colour = "white"
    )
  )
  ) +
  scale_y_continuous(limits = pct_limits,  # EDITED LINE (was: c(-20, 50))
                     breaks = pct_breaks,  # EDITED LINE (was: seq(-20, 60, 10))
                     expand = expansion(mult = c(0.02, 0.02))) +
  labs(x = NULL, y = "Reduction (%) in depression symptom disparity", fill = NULL) +
  theme_classic(base_size = 20) +
  theme(
    panel.grid.major.x = element_line(color = "grey85", size = 0.35),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    strip.background   = element_blank(),
    strip.text         = element_text(size = 14, face = "bold"),
    axis.text          = element_text(color = "black"),
    axis.title.x       = element_text(size = 16, margin = margin(t = 6)),
    legend.position    = "bottom",
    legend.direction   = "horizontal",
    legend.key.height  = unit(10, "pt"),
    legend.key.width   = unit(20, "pt"),
    legend.text        = element_text(size = 15),
    plot.margin        = margin(6, 8, 6, 6)
  ) +
  guides(
  fill = guide_legend(
    order = 1,
    nrow = 2, byrow = TRUE,
    override.aes = list(
      pattern = "none",
      pattern_c_allolour = NA,
      pattern_fill   = NA
    )
  )
)

# EDITED LINE (output path)
ggsave(paste0(plots_dir, "Figure_3_PA.png"), p_c_all, width = 16, height = 8, units = "in", dpi = 300)
```
